---
title: "Broadband access by Texas counties"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

This lesson notes how to use Kyle Walker's [tidycensus](https://walker-data.com/tidycensus/) package to make maps using American Community Survey data via the Census Bureau's API. See [Reporting with Data in R](https://utdata.github.io/rwdir/) for more information.

As of this writing, the [RWDIR chapter is here](https://utdata.github.io/rwdir/census-intro.html).


```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(tidycensus)
library(scales)
```

## Get broadband census data

### Find the data in the portal

The first challenge is to find the variable we need: Broadband any type.

[This is the table](https://data.census.gov/cedsci/table?tid=ACSST5Y2019.S2801).

```{r get-variable}
v19_subject <- load_variables(2019, "acs5/subject", cache = TRUE)

# View(v19_subject)
```


The variable we need: S2801_C02_014

We'll check these numbers to make sure we have the right variables.

### Get our data from tidycensus

Percetage of housholds with broadband

```{r test-variable}
get_acs(
  year = 2019,
  variables = c(broadband = "S2801_C02_014"),
  geography = "us",
)
```

Now that we've confirmed, get for counties in Texas.

```{r bb-counties}
tx_broadband <- get_acs(
  year = 2019,
  variables = "S2801_C02_014",
  geography = "county",
  state = "TX"
)

tx_broadband
```

### Counties with lowest percentage

```{r bb-lowest}
tx_broadband %>% 
  arrange(estimate) %>% 
  head()
```

### Counties with highest percentage

```{r bb-highest}
tx_broadband %>% 
  arrange(estimate %>% desc()) %>% 
  head()
```

## Make the map

### Add the geometry

The same get_acs call, but we are adding geometry.

The output us suppressed.

```{r bb-counties-geo, echo=T, results='hide', message=F, warning=F}
broadband_tx_geo <- get_acs(
  year = 2019,
  variables = "S2801_C02_014",
  geography = "county",
  state = "TX",
  geometry = TRUE
)

```

### Plot the map

```{r bb-plot}
ggplot(broadband_tx_geo) +
  geom_sf(aes(fill = estimate), color = "white", size = .1) +
  theme_void() +
  scale_fill_distiller(
    palette = "Blues",
    direction = 1,
    name = "% Broadband"
  ) +
  labs(
    title = "Broadband connectivity",
    subtitle = "A look at percentage of households with broadband of any type in Texas counties.",
    caption = "Source: American Community Survey/2019 5-year survey"
  )
```


