---
title: "COVID deaths per population"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    df_print: paged
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })
---

This lesson uses COVID case data from the New York Times and Kyle Walker's [tidycensus](https://walker-data.com/tidycensus/index.html) package to make a map of COVID deaths per popluation for Texas counties. See [Reporting with Data in R] for more information.

As of this writing, the [RWDIR chapter for this is here](https://utdata.github.io/rwdir/census-more.html).

## Setup

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(janitor)
library(tidycensus)
```

## Import covid data

The data is found [here](https://github.com/nytimes/covid-19-data).

```{r import-covid}
covid <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv") %>% 
  clean_names()

covid %>% head(10)
```

Get most recent Texas numbers.

There is an "Unknown" county, which we remove.

```{r tx-covid}
tx_covid <- covid %>% 
  filter(state == "Texas") %>% 
  group_by(county) %>% 
  slice_max(date) %>% 
  select(-state) %>% 
  filter(county != "Unknown")

tx_covid
```

## Get population data

First we have to figure out the variable we want. We are using 2020 Decennial data from the redistricting file. This is [the table](https://data.census.gov/cedsci/table?q=P1&tid=DECENNIALPL2020.P1) from data.census.gov that has the population info.

Download the variables from the "pl" table:

```{r load-vars}
vpl <- load_variables(2020, "pl", cache = T)
# view is commented out
# View(vpl)
```

The total population variable is: P1_001N

Test the variable against the portal:

```{r test-var}
get_decennial(
  year = 2020,
  variables = c(total_pop = "P1_001N"),
  geography = "state",
  state = "AL"
)
```

Now let's get Texas by county. The output is suppressed.

```{r tx-pops, echo=T, results='hide', message=F, warning=F}
tx_pop <- get_decennial(
  year = 2020,
  variables = c(total_pop = "P1_001N"),
  geography = "county",
  state = "TX",
  geometry = TRUE # adds shapes
) %>% clean_names()
```

## Join the data

We use an inner join to bring the covid data into the population data.

> You have to start with the population data for the geometry to work.

```{r}
tx_joined <- tx_pop %>% 
  # this is the join line
  inner_join(tx_covid, by = c("geoid" = "fips"))

tx_joined %>% glimpse()
```

### Clean up the columns

Removing columns and renaming value.

```{r join-reselect}
tx_selected <- tx_joined %>% 
  select(-c(date, name, variable)) %>% 
  rename(
    total_pop = value
  )

tx_selected %>% glimpse()
```

### Creates our rate column

Creating cases/deaths per 1,000 population.

```{r create-rates}
tx_rates <- tx_selected %>% 
  mutate(
    cases_per_pop = (cases / (total_pop / 1000)) %>% round(),
    deaths_per_pop = (deaths / (total_pop / 1000)) %>% round(),
  )

tx_rates %>% glimpse()
```

## Highest rates

###  Counties with highest case rates

```{r top-cases}
tx_rates %>% 
  arrange(cases_per_pop %>% desc()) %>% 
  select(county, total_pop, cases, cases_per_pop,) %>% 
  head(10)
```

### Counties with highest deaths rates

```{r top-deaths}
tx_rates %>% 
  arrange(deaths_per_pop %>% desc()) %>% 
  select(county, total_pop, deaths, deaths_per_pop) %>% 
  head(10)
```

## Map the rates

### Cases per 1000


```{r plot-cases}
ggplot(tx_rates) +
  geom_sf(aes(fill = cases_per_pop), color = "white", size = .2) +
  theme_void() +
  labs(
    title = "COVID cases per 1000 in Texas counties",
    caption = "Source: NYTimes, Census Bureau/2020 Decennial"
  ) +
  scale_fill_distiller(
    palette = "Greens",
    direction = 1,
    name = "Cases per 1000"
  )

```

### Deaths per 1000

```{r plot-deaths}
ggplot(tx_rates) +
  geom_sf(aes(fill = deaths_per_pop), color = "white", size = .1) +
  theme_void() +
  labs(
    title = "COVID deaths per 1,000 people in Texas counties",
    caption = "Source: NYTimes, Census Bureau/2020 Decennial"
  ) +
  scale_fill_distiller(
    palette = "Blues",
    direction = 1,
    name = "Deaths per\n1,000 population\n"
  )

```

## Failed join

You have to start with the tibble that has the geometry. This one tries to start with the covid data and the result doesn't work. The error is:

```r
Error: stat_sf requires the following missing aesthetics: geometry Run `rlang::last_error()` to see where the error occurred.
```

The code is commented since it doesn't work.

```{r join-fail}
# tx_joined_flop <- tx_covid %>%
#   inner_join(tx_pop, by = c("fips" = "geoid"))
# 
# tx_joined_flop %>%
#   ggplot() +
#   geom_sf()
```

