---
title: "Census zips practice"
output: html_notebook
---

This is an **abandoned effort** to gets populations of Travis (or some geo) by zip code that could be used later with MixBev lesson.

HUD ZIP [crosswalk file](https://www.huduser.gov/portal/datasets/usps_crosswalk.html).

```{r setup, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(tidycensus)
library(DT)
library(readxl)
library(janitor)
library(scales)
```

## Steps

- find zips in area
  - decide how to cut them: 
- get pops by zip
- filter pops by area zips
- map it


```{r}
zips_raw <- read_excel(
  "data-raw/ZIP_COUNTY_092021.xlsx",
) %>% clean_names()

zips <- zips_raw %>% 
  mutate(
    res_ratio = round(res_ratio, 3),
    bus_ratio = round(bus_ratio, 3),
    oth_ratio = round(oth_ratio, 3),
    tot_ratio = round(tot_ratio, 3),
  )

zips %>% glimpse()
```

### Find our counties

List of FIPS

- Bastrop: 48021
- Caldwell: 48055
- Hays: 48209
- Travis: 48453
- Williamson: 48491


```{r}
fips_msa <- c("48021", "48055", "48209", "48453", "48491")
fips_travis <- c("48453")
fips_htw <- c("48209", "48453", "48491")

zips_filtered <- zips %>% 
  filter(
    county %in% fips_travis,
    bus_ratio > 0.33
  ) %>% 
  arrange(tot_ratio)

zips_filtered
```

```{r}
zips_selected <- zips_filtered %>% 
  select(zip) %>% .$zip
```


## Get variables

Total pop: B01001_001


```{r}
v19 <- load_variables(2019, "acs5/subject", cache = TRUE)

```


```{r}
pop_zip <- get_acs(
  year = 2019,
  variables = "S0101_C01_001",
  geography = "zcta",
  state = "TX",
  # geometry = T,
) %>% clean_names()
  
pop_zip_geo <- get_acs(
  year = 2019,
  variables = "S0101_C01_001",
  geography = "zcta",
  state = "TX",
  geometry = T,
) %>% clean_names()
 
pop_zip
```

```{r}
pop_zip_filtered <- pop_zip_geo %>% 
  filter(geoid %in% zips_selected)

pop_zip_filtered %>% 
  arrange(estimate %>% desc())
```

```{r}
pop_zip_filtered %>% 
  ggplot() +
  geom_sf(aes(fill = estimate)) +
  theme_void() +
  theme(plot.margin = unit(c(10,10,10,10), "pt")) +
  scale_fill_distiller(
    # palette = "Greens",
    direction = 1,
    name = "Population",
    label = comma
  )
```

- This is [where I found to add the margin back](https://stackoverflow.com/questions/40407498/how-to-remove-margins-in-ggplot2-chart)









